plyranges: a grammar for manipulating genomics data

### 2018-02-12 paper.Rmd
added section on overlaps

The _plyranges_ API exports the six core verbs from the _dplyr_ package
and modifys them for use with simple _Ranges_ and _Genomic Ranges_. The
verb `mutate()` takes a Ranges object and a set of name-value pairs
and generates a new Ranges object that with modified or new metadata
columns or modified core components (start, end, width, seqnames, strand). 
The `filter()` function takes a Ranges object and logical expressions and
restricts Ranges object to where the logical expression evaluates to true.
The `summarise()` function takes a Ranges object and a set of name-value pairs
and aggregates the Ranges according to functions evaluated in the name-value pairs.
As `summarise()` is an aggregation it may break the structure of the of a Ranges
object, hence it returns a DataFrame object. The `select()` function determines
which metadata columns are returned and the order they are returned in. The 
`arrange()` function sorts a Ranges object by named variables.
The `group_by()` function creates an implicit grouping of Ranges object according
to variables in the Ranges object. This modifies the actions of `mutate()`,
`summarise()` and `filter()` (and also the set operations, `reduce_ranges()`
and `disjoin_ranges()`), so they are performed on each partition created by the
grouping.

@@

>>>>

The _plyranges_ API exports the six core verbs from the _dplyr_ package
and modifys them for use with simple _Ranges_ and _Genomic Ranges_. The
verb `mutate()` takes a Ranges object and a set of name-value pairs
and generates a new Ranges object that with modified or new metadata
columns or modified core components (start, end, width, seqnames, strand).
The use of `mutate()` means that a user no longer needs knowledge of the 
accessors of the Ranges object, as they can modify them in place.
The `filter()` function takes a Ranges object and logical expressions and
restricts Ranges object to where the logical expression evaluates to true.
The `summarise()` function takes a Ranges object and a set of name-value pairs
and aggregates the Ranges according to functions evaluated in the name-value pairs.
As `summarise()` is an aggregation it may break the structure of the of a Ranges
object, hence it returns a DataFrame object. The `select()` function determines
which metadata columns are returned and the order they are returned in. The 
`arrange()` function sorts a Ranges object by named variables.
The `group_by()` function creates an implicit grouping of Ranges object according
to variables in the Ranges object. This modifies the actions of `mutate()`,
`summarise()` and `filter()` (and also the set operations, `reduce_ranges()`
and `disjoin_ranges()`), so they are performed on each partition created by the
grouping.

@@

A common operation to perform between two _Ranges_ objects is to find overlaps
or nearest neighbours. The _plyranges_ API recasts these operations as 'joins'
or 'pairing' operations. For overlaps, there are three join operations: 
`join_overlap_intersect()`, `join_overlap_inner()` and `join_overlap_left()`
which are shown in figure (). These operations consider any overlap between
two input ranges and return any corresponding metadata from both Ranges objects
as metadata. The intersect join takes the intersect of the start and end coordinates
of overlapping intervals of the query and subject Ranges 
(for GenomicRanges it also accounts for sequence name), when there is a overlap
the metadata corresponding to the query and subject Ranges are returned. 
Similarly, inner join takes the start and end coordinates of the query Ranges
that overlap the subject Ranges and returns metadata of the overlapping query
and subject Ranges. Finally, the left join performs a left outer join between
the query and subject Ranges, it returns all genomic intervals from the query
ranges, and returns missing values in metadata columns when there is no overlap.
A user may also restrict or group by overlaps with the `filter_by_overlaps()`,
`filter_by_non_overlaps()` and `group_by_overlaps()`. All overlap methods can
be modified with the `within` suffix (which changes the type of overlap 
from 'any' to 'within') or the `directed` suffix (which takes into account
the strand of a GenomicRanges object.).

For nearest neighbours, the `plyranges` API provides `join_nearest()`, 
`join_precede()`, and `join_follow()` functions. These functions are similar
to the overlapping functions, in that they return the query ranges that are
nearest (or precede or follow) the subject ranges and add metadata from
the subject ranges when the query is a nearest neighbour of the subject. Like
the overlap joins, these functions can modified with suffixes to find nearest
neighbours that are left, right, upstream or downstream of the subject.

and returns
